PyUltraLight: A Pseudo-Spectral Solver for Ultralight Dark Matter Dynamics

### 2018-09-24 draft.tex
Shaun's rewrite of my typo

While the user is encouraged validate their choice of timestep through convergence testing on a case-by-case basis, \PyUltraLight does include a default timestep value which is chosen based on maximum fluid velocity arguments. In the fluid interpretation of the Schr{\"o}dinger-Poisson system, we define
\begin{equation}\label{eq:madelung}
    \psi\equiv\sqrt{\rho}e^{i\theta}, \quad \vec{v}\equiv\boldsymbol{\nabla}\theta.
\end{equation}
This then allows us to define a maximum fluid velocity that can be successfully modelled on a simulation grid of a given resolution. If the phase change between two consecutive grid points exceeds $\pi$, the fluid will appear to move backwards across the grid. Hence, we define the maximum velocity as that corresponding to a phase change of $\pi$ between grid points. We then set the default timestep, $\Delta t$, such that fluid travelling at this maximum velocity will traverse one grid space, $\Delta x$, per timestep. That is to say
\begin{equation}\label{eq:timestep}
    \Delta t = \frac{(\Delta x)^2}{\pi}.
\end{equation}
We stress that this is merely a choice of timestep, rather than a strict constraint. The motivation for this particular choice of timestep is that if the user finds that their fluid is approaching velocities at which this timestep is no longer sufficiently small for convergence, they will also be approaching the point at which they need to refine the simulation grid. Therefore, choosing the timestep to be smaller than the above default value offers little advantage. In many cases, this choice of timestep will be overly strict, and the user is free to choose a larger timestep by modifying the 'step\textunderscore factor' parameter in the Jupyter notebook.

>>>>

While the user is encouraged to choose their own timestep and validate it through convergence testing on a case-by-case basis, \PyUltraLight does include a default timestep value which is chosen based on maximum fluid velocity arguments. In the fluid interpretation of the Schr{\"o}dinger-Poisson system, we define
\begin{equation}\label{eq:madelung}
    \psi\equiv\sqrt{\rho}e^{i\theta}, \quad \vec{v}\equiv\boldsymbol{\nabla}\theta.
\end{equation}
This then allows us to define a maximum fluid velocity that can be successfully modelled on a simulation grid of a given resolution. If the phase change between two consecutive grid points exceeds $\pi$, the fluid will appear to move backwards across the grid. Hence, we define the maximum velocity as that corresponding to a phase change of $\pi$ between grid points. We then set the default timestep, $\Delta t$, such that fluid travelling at this maximum velocity will traverse one grid space, $\Delta x$, per timestep. That is to say
\begin{equation}\label{eq:timestep}
    \Delta t = \frac{(\Delta x)^2}{\pi}.
\end{equation}
We stress that this is merely a choice of timestep, rather than a strict constraint. The motivation for this particular choice of timestep is that if the user finds that their fluid is approaching velocities at which this timestep is no longer sufficiently small for convergence, they will also be approaching the point at which they need to refine the simulation grid. Therefore, choosing the timestep to be smaller than the above default value offers little advantage. In many cases, this choice of timestep will be overly strict, and the user is free to choose a larger timestep by modifying the 'step\textunderscore factor' parameter in the Jupyter notebook.
